---
title: "Project 1 JSON"
author: "Dexter McNutt"
date: "Oct 18, 2019"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: no
  pdf_document: default
---

```{r setup, include=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(jsonlite)
library(knitr)
library(tidyverse)
```

# JSON What is it?

JSON is a completely language dependent data interchange format that stands for "JavaScript Object Notation" and it is the most widely used data format for data interchange on the web. It's a great way to store data because it's based on a subset of the JavaScript programming language making it easy to understand.

Info found here: 
https://www.geeksforgeeks.org/parse-json-java/

# Using the jsonlite Package

The `JSONLITE` package is simple and has a pretty straightforward syntax and is one of the most widely used packages in R. It also makes saving a data frame in JSON and turning a data frame into a 'flat' JSON file very easy making it a naturally choice.

# The Data

I found a dataset that lists and ranks, based on counts, the most popular baby names. I found the data as open source on data.gov. The variables are the year of birth (birthyear), gender(sex), ethnicity(ethnicity), childs first name(first_name), count(count), and the rank of the name (rank).


## Reading it in
As you can see the `fromJSON()` function in the jsonlite package does a good job of pulling in the data but a little help is needed seperating out the metadata. Some options the `fromJSON()` function has are txt (a JSON string, URL, or file), simplifyVector/simplifyDataFrame/simplifyMatrix (coerce JSON arrays into a vector, data frame, or matrix respectively), and flatten (automatically flatten nested data frames into a single non-nested data frame) just to name a few.
```{r jsonlite, echo=FALSE, message=FALSE, collapse=TRUE}
library(gdata) # for the trim function

readinnow <- function(link) {

  dataRaw <- fromJSON(txt=link,simplifyDataFrame = TRUE)

  data1 <- dataRaw[['data']]

  data2 <- as.data.frame(data1)
  
  return(data2)

}

babyNames <- readinnow('https://data.cityofnewyork.us/api/views/25th-nujf/rows.json?accessType=DOWNLOAD')

babyNamesfinal=subset(babyNames,select=-c(V1,V2,V3,V4,V5,V6,V7,V8))

names(babyNamesfinal) <- c("birthyear", "sex", "ethnicity","first_name","count","rank")

kable(babyNamesfinal[1:5,], caption="Baby Name Data")
```

## Contingency Tables
Here you can see I had to combine a few levels of the column ethnicity:
```{r jsonlite2, echo=TRUE, collapse=FALSE}
levels(babyNamesfinal$ethnicity)

table(kable(babyNamesfinal$ethnicity))

levels(babyNamesfinal$ethnicity) <- c("ASIAN AND PACI","ASIAN AND PACI","BLACK NON HISP","BLACK NON HISP","HISPANIC","WHITE NON HISP", "WHITE NON HISP")

ftable(kable(babyNamesfinal$ethnicity))

levels(babyNamesfinal$ethnicity)
```

## Bar Charts
So if we look at a highly ranked names ""Olivia"" and ""Jordan"" on the list we can see how popular these names are based on gender:
```{r jsonlite3, echo=TRUE, collapse=FALSE}
sex1 <- babyNamesfinal$sex[babyNamesfinal$first_name=="Olivia"]
sex2 <- babyNamesfinal$sex[babyNamesfinal$first_name=="Jordan"]

# Number of Male and Female uses of the name Olivia
barplot(prop.table(table(sex1)))

# Number of Male and Female uses of the name Jordan
barplot(prop.table(table(sex2)))
```

## Boxplots and Scatterplots
```{r jsonlite4, echo=TRUE, collapse=FALSE}
# Simple Scatterplot
attach(babyNamesfinal)
boxplot(as.numeric(count)~as.numeric(rank), main="Box Plots",
   xlab="Ethnicity", ylab="Count")
```

```{r jsonlite5, echo=TRUE, collapse=FALSE}
# Simple Scatterplot
attach(babyNamesfinal)
plot(count, rank, main="Scatterplot Example",
   xlab="Count", ylab="Rank", pch=19)
```










